A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}

### 2022-01-29 ICLP Paper/samplepaper.tex
intro

\usepackage{multirow}
\usepackage{xspace}
\usepackage{url}

\usepackage{comment}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

In real-world production scheduling, the number of operations to process can easily go
into tens of thousands \citep{coltep19a,kohakamo20a,kotakoscge21a}, which exceeds exact
optimization capacities even of state-of-the-art solvers for Answer Set Programming (ASP),
Mixed Integer Programming (MIP), or Constraint Programming (CP)
\citep{daneshamooz2021mathematical,francescutto2021solving,shi2021solving}.
Hence, more efficient approaches to approximate good-quality schedules 
instead of striving for optimal solutions have attracted wide research interest.
On the one hand, respective methods include greedy and local search techniques such as
dispatching rules \citep{blackstone1982state}, shifting bottleneck \citep{adams1988shifting} and
genetic algorithms \citep{pezzella2008genetic}.
On the other hand, problem decomposition strategies based on a
rolling horizon \citep{singer2001decomposition,liu2008prediction} or
bottleneck operations \citep{zhang2010hybrid,zhai2014decomposition} 
have been proposed to partition large-scale instances into better manageable subproblems,
where no single strategy strictly dominates \citep{ovacik2012decomposition}.

@@

In this paper, we greatly extend our preliminary study \citep{elkgeb20a} 
on problem decomposition into time windows and successive schedule optimization
by means of multi-shot ASP modulo DL solving with \clingodl.
The goal of the decomposition is to split highly complex scheduling tasks into
balanced portions for which partial schedules of good quality can be reliably
found within tight runtime limits, and then be merged into a global solution of
significantly better quality than obtainable in similar runtime 
with single-shot optimization on the full problem.
In this process, problem decomposition must respect the precedence of operations
within their encompassing jobs to guarantee the feasibility of schedules, and as
a secondary objective, the order by time windows of operations sharing a machine should come close to the sequences of operations in (unknown) optimal schedules.
We address computational efficiency as well as solution quality by devising and investigating decomposition strategies regarding the size of
time windows and heuristics to choose their operations.

The contributions of our work going beyond the study in \citep{elkgeb20a} are:
\begin{itemize}
\item In addition to problem decomposition based on the earliest starting times
      of operations, we consider the most total work remaining criterion and
      refinements of both strategies by bottleneck machines.
      We encode static as well as dynamic decomposition variants, the latter taking
      partial schedules for operations of previous time windows into account, by stratified
      ASP programs (without DL constraints).
\item Considering that a decomposition into time windows may be incompatible with
      the optimal sequences of operations sharing a machine, we incorporate
      time window overlapping into the iterative scheduling process to
      offer a chance for revising "decomposition mistakes".
      Moreover, as the makespan objective we apply for the optimization of
      (partial) schedules tolerates unnecessary idle times of machines as long as
      they do not yield a greater scheduling horizon,
      we devise a stratified ASP encoding to postprocess and compress partial
      schedules by reassigning operations to earlier idle slots available on their machines.
\item We experimentally evaluate decomposition strategies varying the number of
      operations per time window as well as multi-shot ASP solving
      augmented with overlapping and compression techniques
      on JSP benchmark sets of several sizes.
      In particular, our experiments demonstrate that successive optimization
      by multi-shot ASP solving leads to substantially better schedules within
      tight runtime limits than global optimization on the full problem,
      where the gap increases with the number of operations to schedule.
\end{itemize}

@@

The rest of this 
paper is organized as follows.
Section~\ref{sec:preliminaries} briefly introduces ASP along with the relevant
extensions of multi-shot solving and DL constraints.
In Section~\ref{sec:problem}, we present our successive optimization approach
and detail the ASP programs encoding problem decomposition or iterative scheduling by
time windows, respectively.
Section~\ref{sec:experiments} provides experimental results on JSP benchmark sets, % of several sizes,
assessing different decomposition strategies as well as the impact of overlapping and
compression techniques.
Conclusions and future work are discussed in Section~\ref{sec:conclusions}.  
% The following section shows the most related work and which decomposition techniques have been introduced in the past, followed by the problem formulation section illustrates the problem in detail. Section $4$ presents our proposed model, which shows the decomposition techniques we used and describes one in more detail. In addition, we present the schedule model using Answer Set Programming. Our model is tested by performing extensive experiments on a set of benchmark instances in Section $5$. Section $6$ summarizes this work and provides some ideas to extend the current work.

@@

\paragraph{Multi-shot ASP solving} \citep{gekakasc17a}
allows for iterative reasoning processes by controlling and interleaving
the grounding and search phases of a stateful ASP system.
For referring to a collection of rules to instantiate,
the input language of \clingo supports \lstinline{#program name(c).}
directives, where \lstinline{name} denotes a \emph{subprogram} comprising
the rules below such a directive and the parameter~\lstinline{c} is a
placeholder for some value, e.g., the current time step in case of a
planning problem, supplied upon instantiating the subprogram.
Moreover, \lstinline{#external h : b}$_1$\lstinline{, }$\dots$\lstinline{, b}$_n$\lstinline{.}
statements are formed similar to rules yet declare an atom~\lstinline{h} as
\emph{external} when the body is satisfied:
such an external atom can be freely set to true or false by means of the
Python interface of \clingo,
so that rules including it in the body can be selectively (de)activated
to direct the search.
%
% Multi-shot solving ASP allows solving logic programs dynamically in sequential order. This can be manipulated using APIs implementation via an imperative programming language. Such programming language is used to control the grounding and solving processes and allows the usage of external variables that could be set to $True$ or $False$ to control some logic rules. The main advantage of \emph{multi-shot} solving is to avoid the re-grounding and exploit conflicts learned over time.
%ASP is a paradigm that deals with continuously changing logic programs. In a single-shot approach, an ASP system takes a logic program, computes answer sets and exits. However, the idea of Multi-shot is to consider evolving grounding and solving processes. \emph{clingo} systems enhances the ASP declaritive language with control capacities.

\paragraph{ASP modulo DL}
integrates DL constraints \citep{cotmal06a}, i.e., expressions written as
\lstinline|&diff{t|$_1$\lstinline| - t|$_2$\lstinline|} <= t|$_3$, in the head of rules.
With the exception of the constant~\lstinline{0}, which denotes the number zero,
the terms \lstinline|t|$_1$ and \lstinline|t|$_2$ represent DL variables that can be
assigned any integer value.
However, the difference \lstinline|t|$_1$\lstinline| - t|$_2$ must not exceed the
integer constant~\lstinline|t|$_3$ if the body of a rule with the DL constraint in the
head is satisfied.
That is, the DL constraints asserted by rules whose body is satisfied restrict the
feasible values for DL variables,
and the \clingodl system extends \clingo by checking the consistency of DL constraints
imposed by an answer set.
If these DL constraints are satisfiable, a canonical assignment of smallest feasible
integer values to DL variables can be determined in polynomial time and is output
together with the answer set.
%
% \emph{clingo}[DL] is an extension of the input language \emph{clingo} by theory atoms representing difference constraints \cite{gebser2016theory,janhunen2017clingo}. Difference constraints are defined by specific constraint atoms of the form $\text{\lstinline{&diff}}\{u-v\} \leq k$ where $u$ and $v$ are terms which are interpreted as integer variables and $k$ is an integer constant. \emph{clingo}[DL] provides the following extension of the normal form \eqref{rule1}: 
% \begin{align*}
% 	\text{\lstinline{&diff}}\{u-v\} \leq k \gets a_1, \ldots, a_m, \sim a_{m+1}, \ldots, \sim a_n.
% \end{align*}
%  This rule shows that whenever the body is $True$ the inequality constraints must be satisfied.

>>>>

\usepackage{multirow}
\usepackage{xspace}
\usepackage{url}
\usepackage{enumitem}
\usepackage{comment}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

In real-world production scheduling, the number of operations to process can easily go
into tens of thousands \citep{coltep19a,kohakamo20a,kotakoscge21a}, which exceeds exact
optimization capacities even of state-of-the-art solvers for Answer Set Programming (ASP),
Mixed Integer Programming (MIP), or Constraint Programming (CP)
\citep{daneshamooz2021mathematical,francescutto2021solving,shi2021solving}.
Hence, more efficient approaches to approximate good-quality schedules 
instead of striving for optimal solutions have attracted broad research interest.
On the one hand, respective methods include greedy and local search techniques such as
dispatching rules \citep{blackstone1982state}, shifting bottleneck \citep{adams1988shifting} and
genetic algorithms \citep{pezzella2008genetic}.
On the other hand, problem decomposition strategies based on a
rolling horizon \citep{singer2001decomposition,liu2008prediction} or
bottleneck operations \citep{zhang2010hybrid,zhai2014decomposition} 
have been proposed to partition large-scale instances into better manageable subproblems,
where no single strategy strictly dominates \citep{ovacik2012decomposition}.

@@

In this paper, we significantly extend our preliminary study \citep{elkgeb20a} 
on problem decomposition into time windows and successive schedule optimization
through multi-shot ASP modulo DL solving with \clingodl.
The goal of the decomposition is to split highly complex scheduling tasks into
balanced portions for which partial schedules of good quality can be reliably
found within tight runtime limits. Then, the partial schedules are merged into a global solution of significantly better quality than obtainable in similar runtime with single-shot optimization on the entire problem.
In this process, problem decomposition must satisfy two criteria: 
\begin{enumerate*}[label=\emph{(\roman*)}]
  \item the precedence of operations within their encompassing jobs to guarantee the feasibility of partial schedules, and 
  \item the ordering by time windows of operations sharing a machine should come close to the sequences of operations in (unknown) optimal schedules.
\end{enumerate*}
We address computational efficiency as well as solution quality by devising and investigating decomposition strategies regarding the size of
time windows and heuristics to choose their operations.

This work goes beyond the study of \cite{elkgeb20a} and makes the following contributions:
\begin{itemize}
\item In addition to problem decomposition based on the earliest starting times
      of operations, we consider the most total work remaining criterion and
      refinements of both strategies by bottleneck machines.
      We encode static as well as dynamic decomposition variants, the latter taking
      partial schedules for operations of previous time windows into account, by stratified
      ASP programs (without DL constraints).
\item Since a decomposition into time windows may be incompatible with
      the optimal sequences of operations sharing a machine, we incorporate
      overlapping time windows into the iterative scheduling process to
      offer a chance for revising ``decomposition mistakes''.
      Moreover, the makespan objective, which we apply to optimize (partial) schedules, tolerates unnecessary idle times of machines as long as they do not yield a greater scheduling horizon. 
      Therefore, we devise a stratified ASP encoding to postprocess and compress partial schedules by reassigning operations to earlier idle slots available on their machines.
\item We experimentally evaluate decomposition strategies varying the number of
      operations per time window, as well as multi-shot ASP, solving
      augmented with overlapping and compression techniques
      on JSP benchmark sets of several sizes.
      In particular, our experiments demonstrate that successive optimization
      by multi-shot ASP solving leads to substantially better schedules within
      tight runtime limits than global optimization on the full problem,
      where the gap increases with the number of operations to schedule.
\end{itemize}

@@

The paper is organized as follows:
Section~\ref{sec:preliminaries} briefly introduces ASP along with the relevant
extensions of multi-shot solving and DL constraints.
In Section~\ref{sec:problem}, we present our successive optimization approach
and detail the ASP programs encoding problem decomposition or iterative scheduling by time windows, respectively.
Section~\ref{sec:experiments} provides experimental results on JSP benchmark sets, assessing different decomposition strategies as well as the impact of overlapping and compression techniques.
Conclusions and future work are discussed in Section~\ref{sec:conclusions}.  
% The following section shows the most related work and which decomposition techniques have been introduced in the past, followed by the problem formulation section illustrates the problem in detail. Section $4$ presents our proposed model, which shows the decomposition techniques we used and describes one in more detail. In addition, we present the schedule model using Answer Set Programming. Our model is tested by performing extensive experiments on a set of benchmark instances in Section $5$. Section $6$ summarizes this work and provides some ideas to extend the current work.

@@

\paragraph{Multi-shot ASP solving} \citep{gekakasc17a}
allows for iterative reasoning processes by controlling and interleaving
the grounding and search phases of a stateful ASP system.
For identifying a collection of rules to instantiate,
the input language of \clingo supports \lstinline{#program name(c).}
directives, where the constant~\lstinline{c} is a placeholder for some value, e.g.,
the current time step in case of a planning problem,
supplied upon instantiating the rules below such a directive.

Multi-shot solving ASP allows solving logic programs dynamically in sequential order. This can be manipulated using APIs implementation via an imperative programming language. Such programming language is used to control the grounding and solving processes and allows the usage of external variables that could be set to $True$ or $False$ to control some logic rules. The main advantage of \emph{multi-shot} solving is to avoid the re-grounding and exploit conflicts learned over time.
%ASP is a paradigm that deals with continuously changing logic programs. In a single-shot approach, an ASP system takes a logic program, computes answer sets and exits. However, the idea of Multi-shot is to consider evolving grounding and solving processes. \emph{clingo} systems enhances the ASP declaritive language with control capacities.

\paragraph{ASP Difference Logic:}
\emph{clingo}[DL] is an extension of the input language \emph{clingo} by theory atoms representing difference constraints \cite{gebser2016theory,janhunen2017clingo}. Difference constraints are defined by specific constraint atoms of the form $\text{\lstinline{&diff}}\{u-v\} \leq k$ where $u$ and $v$ are terms which are interpreted as integer variables and $k$ is an integer constant. \emph{clingo}[DL] provides the following extension of the normal form \eqref{rule1}: 
\begin{align*}
	\text{\lstinline{&diff}}\{u-v\} \leq k \gets a_1, \ldots, a_m, \sim a_{m+1}, \ldots, \sim a_n.
\end{align*}
 This rule shows that whenever the body is $True$ the inequality constraints must be satisfied.
